Practical Parametricity for GADTs

### 2020-09-30 20-gadts/inf-rules.tex
set sem for Lan elim added

Finally, to make sense of the final clause
Definition~\ref{def:set-sem}, we first observe that, for any $\rho :
\setenv$, $\textit{Lan}_{\ol{\oCPOsem{\Gamma; \ol\alpha \vdash
      K}\rho[\ol{\alpha := \_}]}} \oCPOsem{\Gamma; \Phi, \ol\alpha
  \vdash F}\rho[\ol{\alpha := \_}]$ is a functor from $\set^m$ to
$\set$ for some $m$. This ensures that
$(\textit{Lan}_{\ol{\oCPOsem{\Gamma; \ol\alpha \vdash
      K}\rho[\ol{\alpha := \_}]}} \oCPOsem{\Gamma; \Phi, \ol\alpha
  \vdash F}\rho[\ol{\alpha := \_}]) \ol{\oCPOsem{\Gamma; \Phi \vdash
    A}\rho}$ is a set.  Moreover, for any $\rho : \setenv$, the fact
that $\oCPOsem{\Gamma;\Phi \vdash (\Lan^{\ol{\alpha}}_{\ol{K}}
  F)\ol{A}}\rho$ is an $\omega$-cocontinuous functor follows from
Corollary~12 of~\cite{jp19} provided each functor in
$\ol{\setsem{\Gamma; \ol\alpha \vdash K}}$ is polynomial in the
variables in $\ol\alpha$. We also note that, for each vector of
functors $\ol K$, $\textit{Lan}_{\ol K}$ is itself a (higher-order)
functor. Specifically, given functors $F, F' : \C \to \D$, a sequence
of functors $\ol K = K_1, ..., K_n$ with $K_i : \C \to \C_i$ for $i =
1,...,n$, and a natural transformation $\alpha : F \to F'$, the
functorial action $\textit{Lan}_{\ol K} \alpha : \textit{Lan}_{\ol K}
F \to \textit{Lan}_{\ol K} F'$ of $\mathit{Lan}_{\ol K}$ on $\alpha$
is defined to be the unique natural transformation such that
$((\textit{Lan}_{\ol K} \alpha) \,\circ \,\langle K_1,...,K_n \rangle)
\circ \eta_F = \eta_{F'} \,\circ\, \alpha$. Here, $\eta_F : F \to
(\textit{Lan}_{\ol K} F) \circ \langle K_1,...,K_n \rangle$ and
$\eta_{F'} : F' \to (\textit{Lan}_{\ol K} F') \circ \langle
K_1,...,K_n \rangle$ are the natural transformations associated with
the functors $\textit{Lan}_{\ol K} F$ and $\textit{Lan}_{\ol K} F'$
from $\Pi_{i \in \{1,...,n\}} \C_i$ to $\D$, respectively. In $\set$,
we have the slightly more general property
\begin{equation}\label{eq:ho-functorial-action-lan}
(\textit{Lan}_{\ol K}\, \alpha)\ol B \circ 
  \iota_{X,g} = \iota'_{X,g} \circ \alpha_X
\end{equation}
Here, $\iota_{X,g}$ is the injection of $FX$ into $\colim{X :
  \set_0,\,g : KX \to B}{FX}$ and $\iota'_{X,g}$ is the injection of
$F'X$ into $\colim{X : \set_0, \,g : KX \to B}{F'X}$.

@@

The next two lemmas show that $ \lambda \ol{N} \,
\ol{B}. (\mu_{\ol{N}})_{\ol{B}} |_{\setsem{\Gamma; \Phi, \ol\beta
    \vdash (\Lan^{\ol\alpha}_{\ol K} F)\ol \beta } \rho[\ol{\Phi,
      \beta := N, B}]}$ is natural in both $\ol{N}$ and $\ol{B}$ and
that it satisfies the condition required for it to be in
$\setsem{\Gamma; \emptyset \vdash \Nat^{\Phi, \ol\beta}\,
  (\Lan^{\ol\alpha}_{\ol K} F)\ol \beta\; G} \rho$, namely, if $\Phi =
{\phi_1,...,\phi_n}$ and $\phi_i$ has arity $k_i$ for $i = 1,...,n$,
then 
\begin{equation}\label{eq:P2}
    \begin{split}
      &\forall \ol{N = (N^1, N^2, N^*) : RT_{k_i}}, \forall \ol{B =
        (B^1, B^2, B^*) :
        RT_0}.\\ &((\mu_{\ol{N^1}})_{\ol{B^1}}|_{\setsem{\Gamma; \Phi,
          \ol\beta \vdash (\Lan^{\ol\beta}_{\ol K} F)\ol \beta }
        \rho[\Phi, \ol \beta := \ol{N^1}, \ol{B^1}]} ,
      (\mu_{\ol{N^2}})_{\ol{B^2}}|_{\setsem{\Gamma; \Phi, \ol\beta
          \vdash (\Lan^{\ol\beta}_{\ol K} F)\ol \beta } \rho[\Phi, \ol
          \beta := \ol{N^2}, \ol{B^2}]} )\\ &: \relsem{\Gamma; \Phi,
        \ol\beta \vdash (\Lan^{\ol\beta}_{\ol K} F)\ol \beta
      }\Eq_{\rho}[\Phi, \ol \beta := \ol N, \ol B] \rightarrow
      \relsem{\Gamma; \Phi, \ol\beta \vdash G}\Eq_{\rho}[\Phi, \ol
        \beta := \ol N, \ol B]
    \end{split}
    \end{equation}

@@

  {\color{red} START HERE!}

To finish the proof, we would like to use that fact that
$\relsem{\Gamma; \Phi,\ol\beta \vdash G}\id_{\Eq_\rho} [\Phi,
   \ol \beta := \ol{\id_N}, \ol{f}]$ is a morphism in $\rel$ and thus preserves
related elements. But in order to apply this morphism, we must first show
that
\begin{align*}
((\mu_{\ol{N^1}})_{\ol{\setsem{\Gamma; \ol\alpha \vdash K} \rho[\ol{\alpha := Z}]}} 
  ((\epsilon_{\ol{N^1}})_{\ol{Z}} \,\, t_1') &, 
  %%%
  (\mu_{\ol{N^2}})_{\ol{\setsem{\Gamma; \ol\alpha \vdash K} \rho[\ol{\alpha := Z}]}} 
  ((\epsilon_{\ol{N^2}})_{\ol{Z}} \,\, t_2')) \\
  & \in 
  \relsem{\Gamma; \Phi, \ol\alpha \vdash
        G[\ol{\beta := K}]}\Eq_{\rho}[\Phi, \ol \alpha := \ol N,
    \ol{\Eq_Z}]
\end{align*}
To show this, we use the condition imposed on $\setsem{\Gamma;
  \emptyset \,|\, \Delta \vdash \eta : \Nat^{\Phi, \ol\alpha} \, F
  \,\, G[\ol{\beta := K}]} \rho \, d$ to be in $\setsem{\Gamma;
  \emptyset \vdash \Nat^{\Phi, \ol\alpha} \, F \,\, G[\ol{\beta :=
      K}]} \rho$, i.e.,

    \begin{equation}\label{eq:P1}
    \begin{split}
      &\forall \ol{N = (N^1, N^2, N^*) : RT_k}, 
          \forall \ol{M = (M^1, M^2, M^*) : RT_k}.\\
      &((\setsem{\Gamma; \emptyset \,|\, \Delta \vdash \eta : \Nat^{\Phi, \ol\alpha} \, F \,\, G[\ol{\beta := K}]} \rho \, d)_{\ol{N^1}, \ol{M^1}}, 
      (\setsem{\Gamma; \emptyset \,|\, \Delta \vdash \eta} : \Nat^{\Phi, \ol\alpha} \, F \,\, G[\ol{\beta := K}] \rho \, d)_{\ol{N^2}, \ol{M^2}}) \\
      &: \relsem{\Gamma; \Phi, \ol\alpha \vdash F}\Eq_{\rho}[\ol{\Phi, \alpha := N, M}]
      \rightarrow \relsem{\Gamma; \Phi, \ol\alpha \vdash
        G[\ol{\beta := K}]}\Eq_{\rho}[\ol{\Phi, \alpha := N, M}]
    \end{split}
    \end{equation}
along with the universal property of 
$\Lan_{\setsem{\Gamma; \ol\alpha \vdash K}\rho {[\ol{\alpha := \_} ]}}
      \setsem{\Gamma; \Phi, \ol\alpha \vdash F}\rho {[\ol{\Phi, \alpha := N^i, \_} ]}$
      for $i = 1, 2$
and the fact that $t_1'$ and $t_2'$ are related in 
$\relsem{\Gamma; \Phi, \ol\alpha \vdash F}\Eq_{\rho}[\ol{\Phi, \alpha := N, \Eq_Z}]$.
Applying Equation \eqref{eq:P1} with $\ol{M} = \ol{Z}$ 
gives (eliding the type of $\eta$)
    \begin{equation*}
    \begin{split}
      &((\setsem{\Gamma; \emptyset \,|\, \Delta \vdash \eta} \rho \, d)_{\ol{N^1}, \ol{Z}}, 
      (\setsem{\Gamma; \emptyset \,|\, \Delta \vdash \eta} \rho \, d)_{\ol{N^2}, \ol{Z}}) \\
      &: \relsem{\Gamma; \Phi, \ol\alpha \vdash F}\Eq_{\rho}[\ol{\Phi, \alpha := N, \Eq_Z}]
      \rightarrow \relsem{\Gamma; \Phi, \ol\alpha \vdash
        G[\ol{\beta := K}]}\Eq_{\rho}[\ol{\Phi, \alpha := N, \Eq_Z}]
    \end{split}
    \end{equation*}
and since we know 
$(t_1',t_2') \in
  \relsem{\Gamma; \Phi, \ol\alpha \vdash F}\Eq_\rho[\ol{\Phi := N}][\ol{\alpha := \Eq_Z}]$, 
we can apply this morphism
%%
to $(t_1', t_2')$:
\begin{align*}
  ((\setsem{\Gamma; \emptyset \,|\, \Delta \vdash \eta} \rho \, d)_{\ol{N^1}, \ol{Z}} t_1', 
  (\setsem{\Gamma; \emptyset \,|\, \Delta \vdash \eta} \rho \, d)_{\ol{N^2}, \ol{Z}} t_2') 
  \in 
\relsem{\Gamma; \Phi, \ol\alpha \vdash
        G[\ol{\beta := K}]}\Eq_{\rho}[\ol{\Phi, \alpha := N, \Eq_Z}]
\end{align*}
%%
Applying the universal property at component $\ol{Z}$
$$
  ((\setsem{\Gamma; \emptyset \,|\, \Delta \vdash \eta}\rho \, d)_{\ol{N^i}})_{\ol{Z}}
  = (\mu_{\ol{N^i}})_{\ol{\setsem{\Gamma; \ol\alpha \vdash K} \rho[\ol{\alpha := Z}]}} \circ 
  (\epsilon_{\ol{N^i}})_{\ol{Z}}
$$
yields 
\begin{align*}
((\mu_{\ol{N^1}})_{\ol{\setsem{\Gamma; \ol\alpha \vdash K} \rho[\ol{\alpha := Z}]}} 
  ((\epsilon_{\ol{N^1}})_{\ol{Z}} \,\, t_1') &, 
  %%%
  (\mu_{\ol{N^2}})_{\ol{\setsem{\Gamma; \ol\alpha \vdash K} \rho[\ol{\alpha := Z}]}} 
  ((\epsilon_{\ol{N^2}})_{\ol{Z}} \,\, t_2')) \\
  & \in 
  \relsem{\Gamma; \Phi, \ol\alpha \vdash
        G[\ol{\beta := K}]}\Eq_{\rho}[\ol{\Phi, \alpha := N, \Eq_Z}] \\
  & = \relsem{\Gamma; \Phi,\ol\beta \vdash G}\Eq_\rho
  [\ol{\Phi, \beta := N, \relsem{\Gamma; \ol\alpha \vdash K}\Eq_\rho[\ol{\alpha := \Eq_Z}]}]
\end{align*}

Finally, we can apply the morphism 
\begin{align*}
  \relsem{\Gamma; \Phi,\ol\beta \vdash G}\id_{\Eq_\rho} [\ol{\Phi, \beta := \id_N, f}] 
  : \,\,
  &\relsem{\Gamma; \Phi,\ol\beta \vdash G}\Eq_\rho
  [\ol{\Phi, \beta := N, \relsem{\Gamma; \ol\alpha \vdash K}\Eq_\rho[\ol{\alpha := \Eq_Z}]}] \\
  \rightarrow \,\,
  &\relsem{\Gamma; \Phi,\ol\beta \vdash G}\Eq_\rho [\ol{\Phi, \beta := N, B}]
\end{align*}
to the pair
$((\mu_{\ol{N^1}})_{\ol{\setsem{\Gamma; \ol\alpha \vdash K} \rho[\ol{\alpha := Z}]}} 
  ((\epsilon_{\ol{N^1}})_{\ol{Z}} \,\, t_1') , 
  %%%
  (\mu_{\ol{N^2}})_{\ol{\setsem{\Gamma; \ol\alpha \vdash K} \rho[\ol{\alpha := Z}]}} 
  ((\epsilon_{\ol{N^2}})_{\ol{Z}} \,\, t_2'))$
, which gives

>>>>

Finally, to make sense of the final clause
Definition~\ref{def:set-sem}, we first observe that, for any $\rho :
\setenv$, $\textit{Lan}_{\ol{\oCPOsem{\Gamma; \ol\alpha \vdash
      K}\rho[\ol{\alpha := \_}]}} \oCPOsem{\Gamma; \Phi, \ol\alpha
  \vdash F}\rho[\ol{\alpha := \_}]$ is a functor from $\set^m$ to
$\set$ for some $m$. This ensures that
$(\textit{Lan}_{\ol{\oCPOsem{\Gamma; \ol\alpha \vdash
      K}\rho[\ol{\alpha := \_}]}} \oCPOsem{\Gamma; \Phi, \ol\alpha
  \vdash F}\rho[\ol{\alpha := \_}]) \ol{\oCPOsem{\Gamma; \Phi \vdash
    A}\rho}$ is a set.  Moreover, for any $\rho : \setenv$, the fact
that $\oCPOsem{\Gamma;\Phi \vdash (\Lan^{\ol{\alpha}}_{\ol{K}}
  F)\ol{A}}\rho$ is an $\omega$-cocontinuous functor follows from
Corollary~12 of~\cite{jp19} provided each functor in
$\ol{\setsem{\Gamma; \ol\alpha \vdash K}}$ is polynomial in the
variables in $\ol\alpha$. We also note that, for each vector of
functors $\ol K$, $\textit{Lan}_{\ol K}$ is itself a (higher-order)
functor. Specifically, given functors $F, F' : \C \to \D$, a sequence
of functors $\ol K = K_1, ..., K_n$ with $K_i : \C \to \C_i$ for $i =
1,...,n$, and a natural transformation $\alpha : F \to F'$, the
functorial action $\textit{Lan}_{\ol K} \alpha : \textit{Lan}_{\ol K}
F \to \textit{Lan}_{\ol K} F'$ of $\mathit{Lan}_{\ol K}$ on $\alpha$
is defined to be the unique natural transformation such that
\begin{equation}\label{eq:functorial-action-of-lan}
((\textit{Lan}_{\ol K} \alpha) \,\circ \,\langle K_1,...,K_n \rangle)
  \circ \eta_F = \eta_{F'} \,\circ\, \alpha
\end{equation}
\noindent
Here, $\eta_F : F \to (\textit{Lan}_{\ol K} F) \circ \langle
K_1,...,K_n \rangle$ and $\eta_{F'} : F' \to (\textit{Lan}_{\ol K} F')
\circ \langle K_1,...,K_n \rangle$ are the natural transformations
associated with the functors $\textit{Lan}_{\ol K} F$ and
$\textit{Lan}_{\ol K} F'$ from $\Pi_{i \in \{1,...,n\}} \C_i$ to $\D$,
respectively. In $\set$, we have the slightly more general property
\begin{equation}\label{eq:ho-functorial-action-lan}
(\textit{Lan}_{\ol K}\, \alpha)\ol B \circ 
  \iota_{X,g} = \iota'_{X,g} \circ \alpha_X
\end{equation}
Here, $\iota_{X,g}$ is the injection of $FX$ into $\colim{X :
  \set_0,\,g : KX \to B}{FX}$ and $\iota'_{X,g}$ is the injection of
$F'X$ into $\colim{X : \set_0, \,g : KX \to B}{F'X}$.

@@

The next two lemmas show that $ \lambda \ol{N} \,
\ol{B}. (\mu_{\ol{N}})_{\ol{B}} |_{\setsem{\Gamma; \Phi, \ol\beta
    \vdash (\Lan^{\ol\alpha}_{\ol K} F)\ol \beta } \rho[\Phi, \ol
    \beta := \ol N, \ol B]}$ is natural in both $\ol{N}$ and $\ol{B}$,
as well as that it satisfies the additional condition required for it
to be in $\setsem{\Gamma; \emptyset \vdash \Nat^{\Phi, \ol\beta}\,
  (\Lan^{\ol\alpha}_{\ol K} F)\ol \beta\; G} \rho$, namely, if $\Phi =
{\phi_1,...,\phi_n}$ and $\phi_i$ has arity $k_i$ for $i = 1,...,n$,
then
\begin{equation}\label{eq:P2}
    \begin{split}
      &\forall \ol{N = (N^1, N^2, N^*) : RT_{k_i}}, \forall \ol{B =
        (B^1, B^2, B^*) :
        RT_0}.\\ &((\mu_{\ol{N^1}})_{\ol{B^1}}|_{\setsem{\Gamma; \Phi,
          \ol\beta \vdash (\Lan^{\ol\beta}_{\ol K} F)\ol \beta }
        \rho[\Phi, \ol \beta := \ol{N^1}, \ol{B^1}]} ,
      (\mu_{\ol{N^2}})_{\ol{B^2}}|_{\setsem{\Gamma; \Phi, \ol\beta
          \vdash (\Lan^{\ol\beta}_{\ol K} F)\ol \beta } \rho[\Phi, \ol
          \beta := \ol{N^2}, \ol{B^2}]} )\\ &: \relsem{\Gamma; \Phi,
        \ol\beta \vdash (\Lan^{\ol\beta}_{\ol K} F)\ol \beta
      }\Eq_{\rho}[\Phi, \ol \beta := \ol N, \ol B] \rightarrow
      \relsem{\Gamma; \Phi, \ol\beta \vdash G}\Eq_{\rho}[\Phi, \ol
        \beta := \ol N, \ol B]
    \end{split}
    \end{equation}

\begin{lemma}
For all $\rho : \setenv$ and $d : \setsem{\Gamma; \emptyset \vdash
  \Delta} \rho$, the family of morphisms
\[ \lambda \ol{N} \, \ol{B}. 
    \ol N, \ol B]} \]
is natural in $\ol{N}$ and $\ol{B}$.
To prove that this family of morphisms is natural, it suffices to show
that the family of unrestricted morphisms $\lambda \ol{N} \,
\ol{B}. (\mu_{\ol{N}})_{\ol{B}}$ is natural in $\ol{N}$ and $\ol{B}$.
Since each $\mu_{\ol N}$ is already natural in $\ol B$ we need only
show that $\mu$ is natural in $\ol N$. To this end, let $\ol{h : N
  \to N'}$, and define functors and natural transformations as
follows:
\ol{K} &= \ol{\lambda \ol M. \oCPOsem{\Gamma; \ol\alpha \vdash K}\rho[\ol{\alpha := M}]} \\
F_{\ol N} &= \lambda \ol M. \oCPOsem{\Gamma; \Phi, \ol\alpha \vdash F}\rho[\Phi := \ol N][\ol{\alpha := M}] \\
F_{\ol{N'}} &= \lambda \ol M. \oCPOsem{\Gamma; \Phi, \ol\alpha \vdash F}\rho[\Phi := \ol{N'}][\ol{\alpha := M}] \\
F_{\ol h} &= \lambda \ol M. \oCPOsem{\Gamma; \Phi, \ol\alpha \vdash F}\id_{\rho}[\Phi := \ol{h}][\ol{\alpha := \id_{M}}] \colon F_{\ol{N}} \to F_{\ol{N'}} \\
G_{\ol N} &= \lambda \ol B. \oCPOsem{\Gamma; \Phi, \ol\beta \vdash G}\rho[\Phi := \ol N][\ol{\beta := B}] \\
G_{\ol{N'}} &= \lambda \ol B. \oCPOsem{\Gamma; \Phi, \ol\beta \vdash G}\rho[\Phi := \ol{N'}][\ol{\beta := B}] \\
G_{\ol h} &= \lambda \ol B. \oCPOsem{\Gamma; \Phi, \ol\beta \vdash G}\id_{\rho}[\Phi := \ol{h}][\ol{\beta := \id_{B}}] \colon G_{\ol{N}} \to G_{\ol{N'}} \\
\eta &= \oCPOsem{\Gamma; \emptyset \,|\, \Delta \vdash \eta : \Nat^{\Phi, \ol\alpha} F\; G[\ol{\beta := K}]} \rho\, d
First note that \(\eta\) is natural, so that
\begin{equation}\label{eq:lan-elim-eta-nat}
\eta_{\ol{N'}} \circ F_{\ol h} = G_{\ol h} \ol{K} \circ \eta_{\ol N} :
F_{\ol N} \to G_{\ol{N'}} \ol{K}
The situation is illustrated in the following diagram:
\begin{tikzcd}[column sep = huge, row sep = huge]
\Set^{|\ol \alpha|}
\ar[r, bend left, "{F_{\ol N}}", ""{name=F, below}]
\ar[r, bend right, "{F_{\ol{N'}}}"{name=F'b, below}, ""{name=F'a, above}]
\ar[Rightarrow, from=F, to=F'a, "{F_{\ol{h}}}"']
\ar[dr, bend right, "{\ol K}"']
&\Set \\
&\Set^{|\ol K|}
\ar[Rightarrow, from=F'b, "{\eta_{\ol{N'}}}"']
\ar[Rightarrow, from=F, "{\eta_{\ol N}}" description]
\ar[u, "{G_{\ol N}}"{name=L, description}]
\ar[u, bend right = 90, "{G_{\ol{N'}}}"{name=L', right}]
\ar[Rightarrow, from=L, to=L', "{G_{\ol h}}"]
Now, note that the collection $\{\epsilon_{\ol N}\}$ actually defines
a natural transformation in $\ol N$ since
\begin{equation}\label{eq:lan-elim-epsilon-nat}
	(\Lan_{\ol K} F_{\ol h}) \ol{K} \circ \epsilon_{\ol N} = \epsilon_{\ol{N'}} \circ F_{\ol h}
holds by~\eqref{eq:functorial-action-of-lan}. That is, we have
\begin{tikzcd}[column sep = huge, row sep = huge]
\Set^j
\ar[r, bend left, "{F_{\ol N}}", ""{name=F, below}]
\ar[r, bend right, "{F_{\ol{N'}}}"{name=F'b, below}, ""{name=F'a, above}]
\ar[Rightarrow, from=F, to=F'a, "{F_{\ol{h}}}"']
\ar[dr, bend right, "{\ol K}"']
&\Set \\
&\Set^k
\ar[Rightarrow, from=F'b, "{\epsilon_{\ol{N'}}}"']
\ar[Rightarrow, from=F, "{\epsilon_{\ol N}}" description]
\ar[u, "{\Lan_{\ol K} F_{\ol N}}"{name=L, description}]
\ar[u, bend right = 90, "{\Lan_{\ol K} F_{\ol{N'}}}"{name=L', right}]
\ar[Rightarrow, from=L, to=L', "{\Lan_K F_{\ol h}}"]
Since $\mu_{\ol N} : \Lan_{\ol K}\, F_{\ol N} \to G_{\ol{N}}$ is the unique
natural transformation such that
\begin{equation}\label{eq:lan-elim-mu-univ}
\mu_{\ol N} \ol{K} \circ \epsilon_{\ol N} = \eta_{\ol N}
and since $\mu_{\ol{N'}} : \Lan_{\ol K}\, F_{\ol{N'}} \to G_{\ol{N'}}$ is
the unique natural transformation such that
\begin{equation}\label{eq:lan-elim-mu'-univ}
\mu_{\ol{N'}} \ol{K} \circ \epsilon_{\ol{N'}} = \eta_{\ol{N'}}
we have the following commuting diagrams:
\begin{tikzcd}[column sep = huge, row sep = huge]
\Set^{|\ol \alpha|}
\ar[r, "{F_{\ol N}}"{name=Fa}, ""{name=F, below}]
\ar[dr, "{\ol K}"']
&\Set \\
&\Set^{|\ol K|}
\ar[Rightarrow, from=F, "{\epsilon_{\ol N}}" description]
\ar[u, "{\Lan_{\ol K} F_{\ol N}}"{name=L, description}]
\ar[u, bend right = 90, "{G_{\ol N}}"{name=G, right}, ""'{name=Gr, right}]
\ar[Rightarrow, bend left = 90, from=Fa, to=Gr, "{\eta_{\ol{N}}}"]
\ar[Rightarrow, from=L, to=G, "{\mu_{\ol N}}"]
%%%
\begin{tikzcd}[column sep = huge, row sep = huge]
\Set^{|\ol \alpha|}
\ar[r, "{F_{\ol{N'}}}"{name=Fa}, ""{name=F, below}]
\ar[dr, "{\ol K}"']
&\Set \\
&\Set^{|\ol K|}
\ar[Rightarrow, from=F, "{\epsilon_{\ol{N'}}}" description]
\ar[u, "{\Lan_{\ol K} F_{\ol{N'}}}"{name=L, description}]
\ar[u, bend right = 90, "{G_{\ol{N'}}}"{name=G, right}, ""'{name=Gr, right}]
\ar[Rightarrow, bend left = 90, from=Fa, to=Gr, "{\eta_{\ol{N'}}}"]
\ar[Rightarrow, from=L, to=G, "{\mu_{\ol{N'}}}"]
By the universal property of $\Lan_K \, F_{\ol{N}}$ there exists a
unique $\theta : \Lan_K \,F_{\ol{N}} \to G_{\ol{N'}}$ such that
\[\theta \ol{K} \circ \epsilon_{\ol N} = \eta_{\ol{N'}} \circ F_{\ol h} = G_{\ol h} \ol{K} \circ \eta_{\ol N}
Here, the last equality is by~\eqref{eq:lan-elim-eta-nat}.
But $G_{\ol h} \circ \mu_{\ol N}$ satisfies this universal property because
	(G_{\ol h} \circ \mu_{\ol N}) \ol{K} \circ \epsilon_{\ol N}
	= G_{\ol h}\ol{K} \circ \mu_{\ol N} \ol{K} \circ \epsilon_{\ol N}
	= G_{\ol h}\ol{K} \circ \eta_{\ol N}
where the second equality is by~\eqref{eq:lan-elim-mu-univ}, and 
$\mu_{\ol{N'}} \circ \Lan_{\ol K} F_{\ol{h}}$ satisfies it as well
because
	(\mu_{\ol{N'}} \circ \Lan_{\ol K} F_{\ol{h}}) \ol{K} \circ \epsilon_{\ol N}
	= \mu_{\ol{N'}}\ol{K} \circ \Lan_{\ol K} F_{\ol{h}}\ol{K} \circ \epsilon_{\ol N}
	= \mu_{\ol{N'}}\ol{K} \circ \epsilon_{\ol{N'}} \circ F_{\ol h}
	= \eta_{\ol{N'}} \circ F_{\ol h}
Here, the second equality is by~\eqref{eq:lan-elim-epsilon-nat} and
the third equality is by~\eqref{eq:lan-elim-mu'-univ}. We must
therefore have that $G_{\ol h} \circ \mu_{\ol N} = \mu_{\ol{N'}}
\circ \Lan_{\ol K} F_{\ol{h}}$ as desired.

@@

To finish the proof, we would like to use that fact that
$\relsem{\Gamma; \Phi,\ol\beta \vdash G}\id_{\Eq_\rho} [\Phi,
   \ol \beta := \ol{\id_N}, \ol{f}]$ is a morphism in $\rel$ and thus preserves
related elements. But in order to apply this morphism, we must first show
that
((\mu_{\ol{N^1}})_{\ol{\setsem{\Gamma; \ol\alpha \vdash K} \rho[\ol{\alpha := Z}]}} 
  ((\epsilon_{\ol{N^1}})_{\ol{Z}} \,\, t_1') &, 
  %%%
  (\mu_{\ol{N^2}})_{\ol{\setsem{\Gamma; \ol\alpha \vdash K} \rho[\ol{\alpha := Z}]}} 
  ((\epsilon_{\ol{N^2}})_{\ol{Z}} \,\, t_2'))
is in its domain
&\relsem{\Gamma; \Phi,\ol\beta \vdash G}\Eq_\rho [\Phi, \ol \beta :=
     \ol N, \ol{\relsem{\Gamma; \ol\alpha \vdash K}\Eq_\rho[\ol{\alpha :=
          \Eq_Z}]}]\\
=~~~~&\relsem{\Gamma; \Phi, \ol\alpha \vdash G[\ol{\beta :=
        K}]}\Eq_{\rho}[\Phi, \ol \alpha := \ol N, \ol{\Eq_Z}]
To do so we note that since $\setsem{\Gamma; \emptyset \,|\, \Delta
  \vdash \eta : \Nat^{\Phi, \ol\alpha} \, F \,\, G[\ol{\beta := K}]}
\rho \, d$ is in $\setsem{\Gamma; \emptyset \vdash \Nat^{\Phi,
    \ol\alpha} \, F \,\, G[\ol{\beta := K}]} \rho$ it must satisfy
    \begin{equation}\label{eq:P1}
      &\forall \ol{N = (N^1, N^2, N^*) : RT_{k_i}}, \forall \ol{M =
        (M^1, M^2, M^*) : RT_0}.\\ &((\setsem{\Gamma; \emptyset \,|\,
        \Delta \vdash \eta : \Nat^{\Phi, \ol\alpha} \, F \,\,
        G[\ol{\beta := K}]} \rho \, d)_{\ol{N^1}, \ol{M^1}},
      (\setsem{\Gamma; \emptyset \,|\, \Delta \vdash \eta} :
      \Nat^{\Phi, \ol\alpha} \, F \,\, G[\ol{\beta := K}] \rho \,
      d)_{\ol{N^2}, \ol{M^2}}) \\ :~~~~ & \relsem{\Gamma; \Phi,
        \ol\alpha \vdash F}\Eq_{\rho}[\Phi, \ol \alpha := \ol N, \ol
        M] \rightarrow \relsem{\Gamma; \Phi, \ol\alpha \vdash
        G[\ol{\beta := K}]}\Eq_{\rho}[\Phi, \ol \alpha := \ol N, \ol
        M]
together with the universal properties of $\Lan_{\ol{\setsem{\Gamma;
      \ol\alpha \vdash K}\rho {[\ol{\alpha := \_} ]}}} \setsem{\Gamma;
  \Phi, \ol\alpha \vdash F}\rho {[\Phi, \ol \alpha := \ol{N^i}, \_ ]}$
for $i = 1, 2$, and the fact that $(t_1',t_2') \in \relsem{\Gamma;
  \Phi, \ol\alpha \vdash F}\Eq_{\rho}[\Phi, \ol\alpha := \ol N,
  \ol{\Eq_Z}]$. Indeed, applying~\eqref{eq:P1} with $\ol{M} = \ol{\Eq_Z}$
gives (eliding the type of $\eta$)
    \begin{equation}\label{eq:morph}
      &((\setsem{\Gamma; \emptyset \,|\, \Delta \vdash \eta} \rho \, d)_{\ol{N^1}, \ol{Z}}, 
      (\setsem{\Gamma; \emptyset \,|\, \Delta \vdash \eta} \rho \, d)_{\ol{N^2}, \ol{Z}}) \\
      &: \relsem{\Gamma; \Phi, \ol\alpha \vdash F}\Eq_{\rho}[\Phi, \ol
        \alpha := \ol N, \ol{\Eq_Z}]
      \rightarrow \relsem{\Gamma; \Phi, \ol\alpha \vdash
        G[\ol{\beta := K}]}\Eq_{\rho}[\Phi, \ol \alpha := \ol N, \ol{\Eq_Z}]
Then since $(t_1',t_2') \in \relsem{\Gamma; \Phi, \ol\alpha \vdash
  F}\Eq_\rho[\Phi := \ol{N}][\ol{\alpha := \Eq_Z}]$, we can apply
~\eqref{eq:morph} to it to get
  ((\setsem{\Gamma; \emptyset \,|\, \Delta \vdash \eta} \rho \, d)_{\ol{N^1}, \ol{Z}} t_1', 
  (\setsem{\Gamma; \emptyset \,|\, \Delta \vdash \eta} \rho \, d)_{\ol{N^2}, \ol{Z}} t_2') 
  \in 
\relsem{\Gamma; \Phi, \ol\alpha \vdash
        G[\ol{\beta := K}]}\Eq_{\rho}[\Phi, \ol \alpha := \ol N, \ol{\Eq_Z}]
%%
Applying the universal property~\eqref{eq:defn-of-mu-sub-N} at component $\ol{Z}$ gives
 ((\setsem{\Gamma; \emptyset \,|\, \Delta \vdash \eta}\rho \, d)_{\ol{N^i}})_{\ol{Z}}
 = (\mu_{\ol{N^i}})_{\ol{\setsem{\Gamma; \ol\alpha \vdash K} \rho[\ol{\alpha := Z}]}} \circ 
  (\epsilon_{\ol{N^i}})_{\ol{Z}}
so that
\big((\mu_{\ol{N^1}})_{\ol{\setsem{\Gamma; \ol\alpha \vdash K} \rho[\ol{\alpha := Z}]}} 
  ((\epsilon_{\ol{N^1}})_{\ol{Z}} \,\, t_1') &, 
  %%%
  (\mu_{\ol{N^2}})_{\ol{\setsem{\Gamma; \ol\alpha \vdash K} \rho[\ol{\alpha := Z}]}} 
  ((\epsilon_{\ol{N^2}})_{\ol{Z}} \,\, t_2')\big) \\
  & \in 
  \relsem{\Gamma; \Phi, \ol\alpha \vdash
        G[\ol{\beta := K}]}\Eq_{\rho}[\Phi, \ol \alpha := \ol N,\ol{\Eq_Z}] 
%  & = \relsem{\Gamma; \Phi,\ol\beta \vdash G}\Eq_\rho
%  [\ol{\Phi, \beta := N, \relsem{\Gamma; \ol\alpha \vdash K}\Eq_\rho[\ol{\alpha := \Eq_Z}]}]
Finally, then, we can apply the morphism 
  \relsem{\Gamma; \Phi,\ol\beta \vdash G}\id_{\Eq_\rho} [\Phi, \ol
    \beta := \ol{\id_N}, \ol f] 
  : \,\,
  &\relsem{\Gamma; \Phi,\ol\beta \vdash G}\Eq_\rho
  [\Phi, \ol \beta := \ol N, \ol{\relsem{\Gamma; \ol\alpha \vdash K}\Eq_\rho[\ol{\alpha := \Eq_Z}]}] \\
  \rightarrow \,\,
  &\relsem{\Gamma; \Phi,\ol\beta \vdash G}\Eq_\rho [\Phi, \ol \beta :=
    \ol N, \ol B]
to the pair
$((\mu_{\ol{N^1}})_{\ol{\setsem{\Gamma; \ol\alpha \vdash K} \rho[\ol{\alpha := Z}]}} 
  ((\epsilon_{\ol{N^1}})_{\ol{Z}} \,\, t_1') , 
  %%%
  (\mu_{\ol{N^2}})_{\ol{\setsem{\Gamma; \ol\alpha \vdash K} \rho[\ol{\alpha := Z}]}} 
  ((\epsilon_{\ol{N^2}})_{\ol{Z}} \,\, t_2'))$.
This gives that
